---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# staggered

<!-- badges: start -->
<!-- badges: end -->

This packages computes the efficient estimator for settings with randomized treatment timing, based on the theoretical results in [Roth and Sant'Anna (2021)](https://arxiv.org/pdf/2102.01291.pdf). If units are randomly (or quasi-randomly) assigned to begin treatment at different dates, the efficient estimator can potentially offer substantial gains over methods that only impose parallel trends. The package also allows for calculating the estimator of [Callaway and Sant'Anna (2020)](https://www.sciencedirect.com/science/article/pii/S0304407620303948?dgcid=author) and the simple-difference-in-means as special cases.

## Installation

You can install the package from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jonathandroth/staggered")
```
## Example

We now illustrate how to use the package by re-creating some of the results in the application section of [Roth and Sant'Anna (2021)](https://arxiv.org/pdf/2102.01291.pdf). Our data contains a balanced panel of police officers in Chicago who were randomly given a procedural justice training on different dates.

### Loading the package and the data

We first load the staggered package as well as some auxiliary packages for modifying and plotting the results.

```{r, echo =T, results = 'hide'}
library(staggered) #load the staggered package
library(dplyr) #load dplyr for data manipulation
library(ggplot2) #load ggplot2 for plotting the results
library(purrr)

df <- staggered::pj_officer_level_balanced #load the officer data
```

We modify the data so that the time dimension is named t, the period of treatment is named g, the outcome is named y, and the individual identifiers are named i.

```{r}
df <- df %>% rename(t = period, y = complaints, g = first_trained, i = uid)
```

### Simple aggregate parameters
We now can call the function calculate_adjusted_estimator_and_se to calculate the efficient estimator. With staggered treatment timing, there are several ways to aggregate treatment effects across cohorts and time periods. The following block of code calculates the simple, calendar-weighted, and cohort-weighted average treatment effects (see Section 3.2 of [Roth and Sant'Anna (2021)](https://arxiv.org/pdf/2102.01291.pdf) for more about different aggregation schemes).


```{r}
#Calculate efficient estimator for the simple weighted average
staggered(df = df, estimand = "simple") %>% select(estimate, se)
```

```{r}
#Calculate efficient estimator for the cohort weighted average
staggered(df = df, estimand = "cohort") %>% select(estimate, se)
```

```{r}
#Calculate efficient estimator for the calendar weighted average
staggered(df = df, estimand = "calendar") %>% select(estimate, se)
```


### Event-study Parameters
We can also calculate an ``event-study'' that computes the average-treatment effect at each lag since treatment.

```{r}
#Calculate event-study coefficients for the first 24 months (month 0 is instantaneous effect)
eventPlotResults <- 
  purrr::map_dfr(.x = 0:23 , #compute event-time effects for first 24 months 
                 .f = ~staggered(
                        df = df, estimand = "eventstudy", eventTime = .x) %>% 
                      mutate(eventTime = .x) )
                   
eventPlotResults %>% select(eventTime, estimate, se) %>% head()
```
```{r}
#Create event-study plot from the results of the event-study
eventPlotResults %>% 
    mutate(ymin_ptwise = estimate + 1.96*se,
           ymax_ptwise = estimate - 1.96*se)%>%
  ggplot(aes(x=eventTime, y =estimate)) +
  geom_pointrange(aes(ymin = ymin_ptwise, ymax = ymax_ptwise))+ 
  geom_hline(yintercept =0) +
  xlab("Event Time") + ylab("Estimate") +
  ggtitle("Effect of Procedural Justice Training on Officer Complaints")
```

### Other Estimators

If instead of the plug-in efficient estimator, one wishes to calculate the simple difference-in-means or Callaway and Sant'Anna estimator, one can specify the argument beta=0 or beta=1, respectively, to the functions above.
